Galaxy Zoo: Morphological Classifications for 150,000 Galaxies in HST Legacy Imaging

### 2016-05-17 writeup/gz_hubble_data.tex

The morphology of galaxies encodes information on the orbital parameters and
assembly history of their contents, including gas, dust, stars, and the central
black hole. The morphology is also closely related to the local environment of
a galaxy, as interactions such as tides, shocks in cluster environments and
direct mergers can all influence a galaxy's internal structure.  For
$M^\star$~galaxies in the local Universe, this interplay between the physical
development of a galaxy and its external appearance typically manifests at the
most basic level as the difference between bulge-dominated, virialized systems
resembling ellipticals (early-types) and disk-dominated, rotationally-supported
systems (late-types) frequently exhibiting spiral arms. This dichotomy has been
used to explore much of the astrophysics governing galaxy formation and
evolution, and has been shown to be closely linked with other galactic
properties such as stellar mass, halo mass, bolometric luminosity, black hole
activity, effective radius, and the relative ages of the stellar populations.

The advent of larger telescopes in an increasing range of observing wavelengths
has revealed that the distribution and properties of galaxy morphology have
strongly evolved over the lifetime of the Universe. At redshift $z\simeq1$
(roughly 6~Gyr after the Big Bang), many galaxies are still in the process of
assembling the baryonic and stellar mass required to reproduce the mature,
coherent structures seen at the present day. This growth occurs via a variety
of pathways, including accretion from large-scale galactic filaments onto halos
via streaming, mergers of individual halos, conversion of gas into stars via
gravitational collapse, etc. The accrual of baryons can be slowed or even
reversed via feedback from stellar winds, supernovae, and active black holes.
Each of these processes affects galaxy morphology in different ways, and so an
accurate measurement of morphological demographics as a function of redshift
provides an extremely powerful observational constraint on the physics involved
\citep[for a recent review see][]{con14}.

Theoretical predictions for the morphology of galaxies as a function of
redshift are primarily computed within the $\Lambda$CDM cosmological framework.
Full treatments model gravitational interactions between baryons and dark
matter, hydrodynamics of the gas, and baryonic physics related to star
formation and evolution. The most advanced simulations now span volumes up to
$\sim100$~Mpc$^3$ while simultaneously resolving the smaller ($<1$ kpc) scales
necessary to reproduce the influence of baryonic physics \citep{vog14a,sch15}.
Such simulations predict clustering of galaxies on large scales in a
hierarchical assembly model \citep{sil12}. The structure of individual galaxies
is affected by their merger history \citep{too72,hop10}, local environment
\citep[such as the morphology-density relation;][]{dre80}, initial dark halo
mass, and many other factors. Morphologies of individual simulated high-mass
galaxies at $z\sim2-3$ commonly show kpc-scale ``clumpy'' structures, with few
galaxies that are either smooth or well-ordered spirals; asymmetric profiles
with strong density contrasts dominate simulated populations in the early
Universe until at least $z\sim1$ \citep{gen14}.

@@

Space-based observatories, particularly the \hubble{} (\hst), have been
responsible for the bulk of imaging studies of high-redshift galaxies.
Observations of fixed fields with very deep imaging
\citep[eg,][]{wil96,gia04,bec06,dav07,sco07,gro11} give the photometric
sensitivity necessary to detect $L^\star$ galaxies at $z>1$, while also
providing the angular resolution to distinguish internal structure and
characterize the morphology. While these measurements are helped by the fact
that the angular diameter distance is relatively flat beyond $z>1$ in a flat
$\Lambda$CDM cosmology, the relevant angular scales are only of the order
$\sim5-10~\mathrm{kpc}/^{\prime\prime}$ \citep{wri06}. \hst{} can thus resolve
much of the structure for a Milky Way-sized galaxy (at least for distinguishing
a disk from a bulge), but will be limited for more compact structures. Since
the size of galaxies evolves as roughly $r\propto(1+z)^{-1}$ \citep{law12a},
the compact sizes of high-redshift galaxies make detailed morphologies a
challenge even for \hst{} \citep{che12}. The public availability of more than
$10^5$~galaxies in archival imaging across various studies gives a data sample
with the potential for high statistical significance.

One of the major difficulties in studying the morphologies of galaxies lies in
the techniques used for measurement. Visual classification by experts has been
used for many decades \citep[eg,][]{hub26,dev59,san61,van76,nai10,bai11,kar15}.
These methods have the advantage of using the significant processing power of
the human brain to identify patterns, but suffer from issues such as lack of
scaling to large surveys and potential issues with replicability and
calibration. Automated measurements, both parametric \citep{pen02a,sim11,lac12}
and non-parametric \citep{abr03,con03,lot04,sca07,bam08,fre13}, scale well to
very large sample sizes, but do not always fully capture the relevant features,
especially for asymmetric galaxies that become increasingly common at high
redshifts. The Galaxy Zoo project \citep{lin08,lin11} utilizes crowdsourced
visual classifications to measure galaxies in color-composite images. The
efforts of more than $2\times10^5$~classifiers allow for multiple independent
classifications of each image which are combined and calibrated to give a
distribution of vote fractions proportional to the probability of a feature
being visible. While crowdsourced data require extensive calibration
\citep{bam09,wil13}, they have a proven reliability and have been used in
dozens of papers
\citep[eg,][]{lan08,bam09,dar10,mas11c,ski12,sim13,sch14,wil15}.

This paper presents the results from the Galaxy Zoo Hubble (GZH)
project.\footnote{\url{http://zoo3.galaxyzoo.org/}} GZH was the third phase of
Galaxy Zoo, following its initial results classifying $\sim900,000$~SDSS~images
into primarily early/late~types \citep{lin11} and Galaxy~Zoo~2, which covered
$\sim250,000$~SDSS~images using a more detailed classification scheme that
included bars, spiral arms, and galactic bulges \citep{wil13}. GZH used a
similarly detailed classification scheme, but focused for the first time on
images of high-redshift galaxies taken with \hst. The Galaxy Zoo: CANDELS
project has also classified morphologies of galaxies at high redshift using ACS
and WFC3 imaging (Simmons et~al., submitted).

>>>>

The morphology of galaxies encodes information on the orbital parameters and
assembly history of their contents, including gas, dust, stars, and the central
black hole. The morphology is also closely related to the local environment of
a galaxy, as interactions such as tides, shocks in cluster environments and
direct mergers can all influence a galaxy's internal structure.  For
$M^\star$~galaxies in the local Universe, this interplay between the physical
development of a galaxy and its external appearance typically manifests at the
most basic level as the difference between bulge-dominated systems with
no/little spiral structure (early-types) and disk-dominated,
rotationally-supported galaxies (late-types) frequently exhibiting spiral arms.
This dichotomy has been used to explore much of the astrophysics governing
galaxy formation and evolution, and has been shown to be closely linked with
other galactic properties such as stellar mass, halo mass, bolometric
luminosity, black hole activity, effective radius, and the relative ages of the
stellar populations.

The advent of larger telescopes in an increasing range of observing wavelengths
has revealed that the distribution and properties of galaxy morphology have
strongly evolved over the lifetime of the Universe. At redshift $z\simeq1$
(roughly 6~Gyr after the Big Bang), many galaxies are still in the process of
assembling the baryonic and stellar mass required to reproduce the mature,
coherent structures seen at the present day. This growth occurs via a variety
of pathways, including accretion from large-scale galactic filaments onto halos
via streaming, mergers of individual halos, conversion of gas into stars via
gravitational collapse, etc. The accrual of baryons can be slowed or even
reversed via feedback from stellar winds, supernovae, and active black holes.
Each of these processes affects galaxy morphology in different ways, and so an
accurate measurement of morphological demographics as a function of redshift
provides an extremely powerful observational constraint on the physics involved
\citep[for recent reviews see][]{but13,con14}.

Theoretical predictions for the morphology of galaxies as a function of
redshift are primarily computed within the $\Lambda$CDM cosmological framework.
Full treatments model gravitational interactions between baryons and dark
matter, hydrodynamics of the gas, and baryonic physics related to star
formation and evolution. The most advanced simulations now span volumes up to
$\sim100$~Mpc$^3$ while simultaneously resolving the smaller ($<1$ kpc) scales
necessary to reproduce the influence of baryonic physics \citep{vog14a,sch15}.
Such simulations predict clustering of galaxies on large scales in a
hierarchical assembly model \citep{sil12}. The structure of individual galaxies
is affected by their merger history \citep{too72,ste02,hop10}, local
environment \citep[such as the morphology-density relation;][]{dre80}, initial
dark halo mass, secular evolution rate, and many other factors. Morphologies of
individual simulated high-mass galaxies at $z\sim2-3$ commonly show kpc-scale
``clumpy'' structures, with few galaxies that are either smooth or well-ordered
spirals; asymmetric profiles with strong density contrasts dominate simulated
populations in the early Universe until at least $z\sim1$ \citep{bel12,gen14}.

@@

Space-based observatories, particularly the \hubble{} (\hst), have been
responsible for the bulk of imaging studies of high-redshift galaxies.
Observations of fixed fields with very deep imaging
\citep[eg,][]{wil96,gia04,bec06,dav07,sco07,gro11} give the photometric
sensitivity necessary to detect $L^\star$ galaxies at $z>1$, while also
providing the angular resolution to distinguish internal structure and
characterize the morphology. While these measurements are helped by the fact
that the angular diameter distance is relatively flat beyond $z>1$ in a flat
$\Lambda$CDM cosmology, the relevant angular scales are only of the order
$\sim5-10~\mathrm{kpc}/^{\prime\prime}$ \citep{wri06}. \hst{} can thus resolve
much of the structure for a Milky Way-sized galaxy (at least for distinguishing
a disk from a bulge), but will be limited for more compact structures. Since
the size of galaxies evolves as roughly $r\propto(1+z)^{-1}$
\citep{mao98,law12a}, the compact sizes of high-redshift galaxies make detailed
morphologies a challenge even for \hst{} \citep{che12}. The public availability
of more than $10^5$~galaxies in archival imaging across various studies gives a
data sample with the potential for high statistical significance.

One of the major difficulties in studying the morphologies of galaxies lies in
the techniques used for measurement. Visual classification by experts has been
used for many decades \citep[eg,][]{hub26,dev59,san61,van76,nai10,bai11,kar15}.
These methods have the advantage of using the significant processing power of
the human brain to identify patterns, but suffer from issues such as lack of
scaling to large surveys and potential issues with replicability and
calibration. Automated measurements, both parametric \citep{pen02a,sim11,lac12}
and non-parametric \citep{abr03,con03,lot04,sca07,bam08,fre13}, scale well to
very large sample sizes, but do not always fully capture the relevant features,
especially for asymmetric galaxies that become increasingly common at high
redshifts. The Galaxy Zoo project \citep{lin08} utilizes crowdsourced visual
classifications to measure galaxies in color-composite images. The efforts of
more than $2\times10^5$~classifiers allow for multiple independent
classifications of each image which are combined and calibrated to give a
distribution of vote fractions proportional to the probability of a feature
being visible. While crowdsourced data require extensive calibration
\citep{bam09,wil13}, they have a proven reliability and have been used for a
wide variety of scientific studies
\citep[eg,][]{lan08,bam09,dar10,mas11c,ski12,sim13,sch14,wil15}.

This paper presents the classifications collected from the Galaxy Zoo Hubble
(GZH) project.\footnote{\url{http://zoo3.galaxyzoo.org/}} GZH was the third
phase of Galaxy Zoo, following its initial results classifying
$\sim900,000$~SDSS~images into primarily early/late~types \citep{lin11} and
Galaxy~Zoo~2, which covered $\sim250,000$~SDSS~images using a more detailed
classification scheme that included bars, spiral arms, and galactic bulges
\citep{wil13}. GZH used a similarly detailed classification scheme, but focused
for the first time on images of high-redshift galaxies taken with \hst. The
Galaxy Zoo: CANDELS project has also classified morphologies of galaxies at
high redshift using ACS and WFC3 imaging (Simmons et~al., submitted).
